Joint training of classification model and similarity model for low-resource text classification in chatbot

### 2020-08-30 chatbot_classification/aaai_2021/versions/paper.ver_01.tex
Reorganized.

Another common approach is text classification neural network based on label-word joint embedding, such as LEAM\cite{wang2018joint}, MTLE\cite{zhang2017multi} and EXAM\cite{du2019explicit}. This approach involves label embedding when constructing the text-sequence representation. Take LEAM as an example, the text classification model is implemented by jointly embedding the word and label in the same latent space\cite{wang2018joint}. Afterwards, the text representation are constructed based on text-label compatibility through attention mechanism. In this way, additional sources of semantic information from class labels can be fully leveraged, which is quite helpful especially under few-shot scenario.

The third popular approach among previous work was based on sentence-pair model. The motivation of this approach was started from the idea of Information Retrival (IR) based chatbot\cite{jafarpour2010filter, leuski2011npceditor}. Having a Q-A (Question-Answer) pairs dataset and user query Q, the IR based conversational system will look up in the Q-A dataset for the pair (Q', A') that best matches query Q through semantic analysis and returns A' as the answer to Q\cite{mnasri2019recent}. In this way, sentence-pair model pre-trained on large corpus of semantic similarity identification task can be applied for being used as a tool to identify the class with highest semantic similarity to customer's query. A common model structure of pre-trained\cite{devlin2018bert} sentence-pair model is based on multiple cross-attention mechanism\cite{barkan2020scalable}, shown in Fig. \ref{Fig.1.sub.2}. Due to the fact that many experiment results have shown that RoBERTa\cite{liu2019roberta} is an improved  version  of  BERT, and has achieved amazing results on both text classification and sentence-pairs semantic similarity tasks, we choose RoBERTa as an important baseline approach in this paper.

Despite the success of these 3 approaches, they still have some limitations in task-specific chatbot scenario. As for the text classification model approach, it's quite hard to use data augmentation method to solve the data insufficiency challenge since the it's usually unfeasible to get large amount of domain-specific data when facing a new task. As for label-word joint embedding approach, it poses noticeable requirements on label information. However, in our cases, since we have large amount of class labels without being well-defined for question answering chatbot, the only thing we can do is to choose one of the user utterance as the class label to represent the key information of the whole class. In this way, the label will become not only prolix but also biased, which will bring some negative effect to label embedding. With respect to the sentence-pair model approach, though we can obtain large amount of data in semantically duplicate sentence pair identification domain for transfer learning\cite{sun2019fine}, it's still quite hard to perform well on task-specific dataset because the training objective of sentence-pair model is semantic similarity, which is slightly different from the target of classification task. That is to say, there always exists some intent classes which cannot be distinguished by each other merely depending on semantic similarity. For example, we have 2 user query saying, A: What should I do if I want to change my password? B: What is the modification rule if I want to change my password? In this case, sentence A is semantically similar to B, since they both express the 
desire for changing password. However, it's still reasonable to classify them into 2 intent classes, since A is asking for the method for changing password, while B is asking for the rule to follow when creating a new password.

@@

\section{Methodology}
In this section, we will describe a system of similarity model fused with classification model, named as SFC, in detail. We come up with 2 different approaches for implementation of SFC system, in which hierarchical SFC is a further improvement of 2-stage SFC for better performance. 
\subsection{2-stage SFC with multi-task learning}
The detailed network structure of 2-stage SFC is shown in Fig. \ref{fig:ver.1.3}. It is composed with 2 separate stages in charge of different jobs.
In stage 1, we can use any auxiliary model or tool to provide top-K most related class labels for sampling sentence pairs of both positive samples and negative samples for stage 2. Suppose we have a training data set $\mathcal{D}=\{x_{i},y_{i}\}_{i=1}^{N}$ of $N$ data points, in which $x_{i}$ is the user input query and $y_{i}$ is a single class label from a class label set of $C$ classes in total. If we want to generate sentence pair data set $\mathcal{D'}=\{[(x,x')_{j},l_{j}]\}_{j=1}^{M}$, where $(x, x')_{j}$ is sentence pair grouped by 2 user input queries and $l_{j}\in\{0, 1\}$ is similarity class label, see Equ. \ref{eq:similarity label}, 
\begin{align}
l_{j}=\begin{cases} \label{eq:similarity label}
0 & y_{x}\neq y_{x'}\\
1 & y_{x}=y_{x'}
\end{cases}
\end{align}
directly from original data set $\mathcal{D}$ without any sampling strategy, then $\mathcal{D'}$ will have $M=N^2$ data points of sentence pairs, which might be too time-consuming for the training process of stage 2 since we often have hundreds of class labels under task-specific chatbot scenario. Although we may only have a few data points for a single class label, $M$ is still quite a big number.

Therefore, we adopt the idea of adversarial negative sampling strategy raised by \cite{bamler2020extreme}. The key idea of this sampling strategy is to train with negative samples that are hard to be distinguished from positive samples. Now suppose we have a user input query $x_{i}$ and top-K most related class labels set $\mathcal{C'}=\{c'_{1}, \dots, c'_{K}\}$ provided by auxiliary model or tool in stage 1, where $K$ is a hyperparameter set by us that controls candidate class number, we apply the adversarial negative sampling strategy into our sentence pair sampling process by the following 2 steps:

\textbf{1. positive sentence pair sampling}: During the training process, we have to first make sure that the ground truth class label $y_{i}$ for $x_{i}$ is within the candidate class label set $\mathcal{C'}$. If $y_{i}\notin \mathcal{C'}$, we will manually add $y_{i}$ into $\mathcal{C'}$ by replacing $c'_{K}$ with $y_{i}$, since $c'_{K}$ is the least promising candidate label according to the auxiliary model in stage 1. Afterwards, we will randomly sample $P$ sentences $\{x'_{0},x'_{1}, \dots, x'_{P}\}$ with the same class label as $y_{i}$ from data set $\mathcal{D}$ to form the set of sentence pairs with positive label $\mathcal{P'}_{i}=\{[(x_{i}, x'_{0}), 1], [(x_{i}, x'_{1}), 1], \dots, [(x_{i}, x'_{P}), 1]\}$, where $P$ is also a hyperparameter set by us that controls the number of sentences we should sample from each class.

\textbf{2. negative sentence pair sampling}: As for negative sentence pairs, we will also randomly sample $P$ sentences for each class in the negative candidate class label set $\mathcal{C'}\backslash \{y_{i}\}$. The class labels in $\mathcal{C'}\backslash \{y_{i}\}$ are the set of most confusing class labels comparing to the ground truth ${y_{i}}$, so we assume that the sentence pairs with negative label grouped by user input query $x_{i}$ and sentences with class labels in $\mathcal{C'}\backslash \{y_{i}\}$ are strong adversarial negative samples for sentence-pair similarity model that can help enhance the training speed and performance. According to the same method as positive sentence pair sampling, we can obtain the set of sentence pairs with negative label as $\mathcal{N'}_{i}=\{[(x_{i}, x'_{0}), 0], [(x_{i}, x'_{1}), 0], ..., [(x_{i}, x'_{P\cdot (K-1)}), 0]\}$

@@

\subsubsection{stage 1: classification model for providing top-K candidate class labels}
As for stage 1, our SFC system can adapt to any auxiliary model or even searching tool based on Term Frequency-Inverse
Document Frequency (TF-IDF) retrieval. In this paper, however, we choose to fine-tune RoBERTa on our task-specific data set $\mathcal{D}$ by adding a softmax classifier on top of it, the same as the structure shown in Fig. \ref{Fig.1.sub.1}, for better overall performance. According to our experimental result, RoBERTa can guarantee that approximately 95\% ground truth label be included in top-5 candidate class labels with highest probability after being fine-tuned on task-specific data set with experimental result measured by F-score at around 0.75. So we use RoBERTa as encoder of classification model in our setting at stage 1 to ensure low limitation for the performance of stage 2.

@@

Before we start to fine-tune our sentence-pair model on task-specific dataset, we first fine-tune RoBERTa on Quora dataset\cite{iyer2017first}, which contains 404,290 potential duplicate question pairs, for transfer learning. Then we continue to do multi-task training on sentence-pair dataset $\mathcal{D'}$ sampled from the top-K candidate class labels provided by stage 1.

@@

As for top-K classification task, it shares the same sentence-pair similarity model structure with task 1. The only difference is that we add a task-specific average pooling layer, shown in Fig. \ref{fig:ver.1.3} to implement the classification task based on sentence-pair model.

We already know that the size of ${\Phi}^S$ in task 1 is $M\times 2=(N\cdot K\cdot P)\times 2$, where $N$ is the total number of data points in original training data $\mathcal{D}$, $K$ is a hyperparameter that controls the number of candidate class labels provided by stage 1, and $P$ is also a hyperparameter that controls the number of sentences we should randomly sample from each candidate class labels. Now, for the average pooling layer, we first reshape ${\Phi}^S$ into the size of $N\times K\times P\times 2$, and then split it into ${\Phi}^{K,pos}$ and ${\Phi}^{K,neg}$, and both of them will have the size of $N\times K\times P$. In this way, ${\Phi}^{K,pos}$ can represent the level of similarity for each sentence pair, and in the mean time, ${\Phi}^{K,neg}$ can represent the level of dissimilarity for each sentence pair. Then, we can do average pooling for each candidate class among the top-K classes as shown in Equ. \ref{eq:average pooling:1} and Equ. \ref{eq:average pooling:2}.

@@

Therefore, to further improve the overall performance of SFC, we proposed a hierarchical joint model structure, shown in Fig. \ref{fig:framework}. In hierarchical SFC, classification model is being placed in lower layer level to dynamically provide top-K candidate class labels for the sentence-pair similarity model placed in higher layer level through a top-K pooling layer. In this way, classification model and similarity model can be merged into one single joint-model for multi-task training with sentence pairs sampled from varying candidate class labels, thus avoiding the limitation brought by 2 separate stage structure.

@@

\section{Experiments}
In this section, we conduct extensive experiments over 5 task-specific datasets to evaluate the performance of our proposed SFC system and the effectiveness of the methodology of multi-task learning and hierachical joint-model structure under low-data settings.

\subsection{Datasets}
The datasets used in our experiments aims to help us study low-resource text classification problem under task-specific conversational chatbot scenario. Therefore, we keep all the dataset under low-resource chatbot corpus setting, which is a large number of class labels (50-250) with a few examples(5-20) in short text for each class, by sampling smaller subsets from the full scale when the raw datasets are oversized, see Table \ref{tbe:dataset statistic}.

@@

$\bullet$ \textbf{Full-scale real-world task-specific faq chatbot dataset}

@@

$\bullet$ \textbf{Subset of customer review dataset for extreme classification}

@@

$\bullet$ \textbf{Subset of recently-released intent detection datasets}.

@@

TextCNN & \multirow{2}*{$0.5318$} & \multirow{2}*{$0.6963$} & \multirow{2}*{$0.7609$} & \multirow{2}*{$0.4408$} & \multirow{2}*{$0.6436$} & \multirow{2}*{$0.7366$} & \multirow{2}*{$0.7918$} & \multirow{2}*{$0.3112$} & \multirow{2}*{$0.4007$} & \multirow{2}*{$0.4823$} & \multirow{2}*{$0.5272$} & \multirow{2}*{$0.6624$} & \multirow{2}*{$0.4401$}\tabularnewline
LEAM & \multirow{2}*{$0.7514$} & \multirow{2}*{$0.8203$} & \multirow{2}*{$0.8612$} & \multirow{2}*{$0.5422$} & \multirow{2}*{$0.7812$} & \multirow{2}*{$0.8129$} & \multirow{2}*{$0.8280$} & \multirow{2}*{$0.4545$} & \multirow{2}*{$0.5554$} & \multirow{2}*{$0.5936$} & \multirow{2}*{$0.6599$} & \multirow{2}*{$0.7086$} & \multirow{2}*{$0.6091$}\tabularnewline

@@

ALBERT-xxlarge & \multirow{2}*{$0.6500$} & \multirow{2}*{$0.9008$} & \multirow{2}*{$0.9296$} & \multirow{2}*{$0.3608$} & \multirow{2}*{$0.6345$} & \multirow{2}*{$0.8090$} & \multirow{2}*{$0.8530$} & \multirow{2}*{$0.3549$} & \multirow{2}*{$0.3876$} & \multirow{2}*{$0.5808$} & \multirow{2}*{$0.6320$} & \multirow{2}*{$0.7253$} & \multirow{2}*{$0.6893$}\tabularnewline

@@

RoBERTa-large & \multirow{2}*{$0.8266$} & \multirow{2}*{$0.8861$} & \multirow{2}*{$0.9023$} & \multirow{2}*{$0.757$} & \multirow{2}*{$0.8476$} & \multirow{2}*{$0.8614$} & \multirow{2}*{$0.8743$} & \multirow{2}*{$0.5425$} & \multirow{2}*{$0.6164$} & \multirow{2}*{$0.6503$} & \multirow{2}*{$0.6729$} & \multirow{2}*{$0.7418$} & \multirow{2}*{$0.6362$}\tabularnewline

@@

2-stage SFC & \multirow{2}*{$0.9162$} & \multirow{2}*{$0.9424$} & \multirow{2}*{$0.9530$} & \multirow{2}*{$0.7997$} & \multirow{2}*{$0.8823$} & \multirow{2}*{$0.8945$} & \multirow{2}*{$0.9123$} & \multirow{2}*{$0.6498$} & \multirow{2}*{$0.6980$} & \multirow{2}*{$0.7202$} & \multirow{2}*{$0.7358$} & \multirow{2}*{$0.8020$} & \multirow{2}*{$0.7311$}\tabularnewline

@@

\textbf{Group 1: single-task classification model.} This group of baseline models are based on the network structure for classification model shown in \ref{Fig.1.sub.1}. As for non-pretrained neural network, we choose TextCNN\cite{kim2014convolutional} as a representative baseline. As for the pretrained model, we choose RoBERTa, BERT and ALBERT, all of which are among the most powerful pretrained models that achieve outstanding results in many NLP tasks, in our experimental setting. Since we apply RoBERTa in our SFC system, we naturally adopt single-task classification model based on RoBERTa-large and RoBERTa-base as our baseline models. Besides, we also adopt the recently released baseline pre-trained model ALBERT-xxlarge\cite{lan2019albert} for comparison. Since we use different few-shot settings from \cite{casanueva2020efficient} for those intent detection datasets, we also use BERT-large tuned on our few-shot settings as an essential reference for experimental results.

@@

\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
\multicolumn{1}{|c|}{\multirow{2}*{Dataset}}&\multicolumn{1}{c|}{\multirow{2}*{Model}}& $K=3$ & $K=5$ & $K=10$ & $K=15$ & $K=20$\tabularnewline
\multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & $P=20$ & $P=10$ & $P=5$ & $P=4$ & $P=3$\tabularnewline
\hline
\multicolumn{1}{|c|}{\multirow{2}*{ITG}}& 2-stage SFC & $0.8034$ & $0.8124$ & $0.8008$ & $0.7967$ & $0.7934$\tabularnewline
\cline{2-7}
\multicolumn{1}{|c|}{} & Hierarchical SFC & $0.7882$ & $0.8114$ & $0.8010$ & $0.7972$ & $0.7826$\tabularnewline
\hline
\multicolumn{1}{|c|}{\multirow{2}*{Amazon-670k}}& 2-stage SFC & $0.7278$ & $0.7364$ & $0.7366$ & $0.7328$ & $0.7204$\tabularnewline
\cline{2-7}
\multicolumn{1}{|c|}{} & Hierarchical SFC & $0.7334$ & $0.7445$ & $0.7516$ & $0.7344$ & $0.7373$\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\caption{We show the performance of SFC from different settings of hyperparameters for $K$, the candidate class number, and $P$, the sample number for each class.}

@@

\textbf{Comparison among SFC variants.} First, we turn to compare among the four SFC variants shown in Table \ref{tbe:table2} to analyze the improvement brought by multi-task and hierarchical joint-model structure. We conduct ablation experiments to study the benefits brought by multi-task training by comparing multi-task 2-stage SFC with single-task 2-stage SFC. We can see from Table \ref{tbe:table2} that 2-stage SFC with only task 1 (sentence-pair similarity task) degrades by 0.65 percent on average and 2-stage SFC with only task 2 (top-K classification task) degrades by 0.3 percent on average. The degradation indicates that training with these 2 tasks together in SFC can effectively help the sentence-pair model extract more knowledge from the limited amount of data. Besides, task 2 degrades 0.35 percent less than task 1. This observation suggests that task 2 may play a relatively more important role in multi-task learning, which leads us to put larger weight $\alpha_{K}$ to the loss $\mathcal{L}^{K}$ of task 2 when calculating the total loss $\mathcal{L}$. Also, we can see that the hierachical SFC steadily outperforms 2-stage SFC on 5 diverse datasets by 1.17 percents on average. This observation supports the idea that the hierachical joint model structure can alleviate the limitation brought by the lack of interaction between similarity model and classification model in 2-stage SFC. Therefore, for the following analysis, we will focus on the comparison between hierachical SFC and the other baseline models.

\textbf{Analysis under few-shot scenarios.} Next, we analyze the overall improvement trend of SFC under various few-shot settings. Our main focus is to study chatbot building in common situation during real-world application where only a few sample sentences are available for each class. The experimental results for the 3 diverse intent detection datasets (ClINC150, BANKING77, HWU64) under few-shot settings indicate that our proposed 
hierarchical SFC can achieve 2.47 percents average improvement over one of the most powerful baseline models, which is RoBERTa-large classification model, on all these few-shot settings. Moreover, the improvement in F-score becomes more and more prominent with the decrease of data sample number per class. Our hierarchical SFC can achieve 4.97 percents average improvements over RoBERTa-large classification model in the most extreme few-shot setting with only 5 training samples per class. This observation indicates that SFC is especially suitable for being applied during the initial stage of building task-specific chatbot system, where the amount of data sample for each class label is extremely scarce.

\textbf{Comparison between label embedding and sentence pair embedding.} Afterwards, we start to compare the performance of SFC with the text-label joint embedding baseline named as LEAM. As shown in Table \ref{tbe:table2}, hierarchical SFC outperforms LEAM by 14.25 percents on average F-score over 5 datasets. Besides, although the labels within the 3 diverse intent detection datasets (CLINC150, BANKING77, HWU64) are all being well-defined as the semantic summary of each intent, hierarchical SFC still outperforms LEAM by 14.69 percents on average F-score. This observation indicates that it's usually quite hard to use a fixed label with only a few words as the representative of all the sentences within a class since there are always plenty of ways for people to express questions or queries about the same topic during a conversation with chatbot. Therefore, learning the semantic similarity between user query and sample sentences through sentence pair embedding can be more reliable than making decision according to the semantic relationship between user query and class labels through label embedding.

>>>>

Another popular approach among previous work was based on sentence-pair model. The motivation of this approach was started from the idea of Information Retrival (IR) based chatbot\cite{jafarpour2010filter, leuski2011npceditor}. Having a Q-A (Question-Answer) pairs dataset and user query Q, the IR based conversational system will look up in the Q-A dataset for the pair (Q', A') that best matches query Q through semantic analysis and returns A' as the answer to Q\cite{mnasri2019recent}. In this way, sentence-pair model pre-trained on large corpus of semantic similarity identification task can be applied for being used as a tool to identify the class with highest semantic similarity to customer's query. A common model structure of pre-trained\cite{devlin2018bert} sentence-pair model is based on multiple cross-attention mechanism\cite{barkan2020scalable}, shown in Fig. \ref{Fig.1.sub.2}. Due to the fact that many experiment results have shown that RoBERTa\cite{liu2019roberta} is an improved  version  of  BERT, and has achieved amazing results on both text classification and sentence-pairs semantic similarity tasks, we choose RoBERTa as an important baseline approach in this paper.

Despite the success of these 2 approaches, they still have some limitations in task-specific chatbot scenario. As for the text classification model approach, it's quite hard to use data augmentation method to solve the data insufficiency challenge since the it's usually unfeasible to get large amount of domain-specific data when facing a new task. With respect to the sentence-pair model approach, though we can obtain large amount of data in semantically duplicate sentence pair identification domain for transfer learning\cite{sun2019fine}, it's still quite hard to perform well on task-specific dataset because the training objective of sentence-pair model is semantic similarity, which is slightly different from the target of classification task. That is to say, there always exists some intent classes which cannot be distinguished by each other merely depending on semantic similarity. For example, we have 2 user query saying, A: What should I do if I want to change my password? B: What is the modification rule if I want to change my password? In this case, sentence A is semantically similar to B, since they both express the 
desire for changing password. However, it's still reasonable to classify them into 2 intent classes, since A is asking for the method for changing password, while B is asking for the rule to follow when creating a new password.

@@

\section{Methodology}
In this section, we will describe a system of similarity model fused with classification model, named as SFC, in detail. We come up with 2 different approaches for implementation of SFC system, in which hierarchical SFC is a further improvement of 2-stage SFC for better performance. 
\subsection{2-stage SFC with multi-task learning}
The detailed network structure of 2-stage SFC is shown in Fig. \ref{fig:ver.1.3}. It is composed with 2 separate stages in charge of different jobs.
In stage 1, we can use any auxiliary model or tool to provide top-K most related classes for sampling sentence pairs of both positive samples and negative samples for stage 2. Suppose we have a training data set $\mathcal{D}=\{x_{i},y_{i}\}_{i=1}^{N}$ of $N$ data points, in which $x_{i}$ is the user input query and $y_{i}$ is a single class label from a class label set of $C$ classes in total. If we want to generate sentence pair data set $\mathcal{D'}=\{[(x,x')_{j},l_{j}]\}_{j=1}^{M}$, where $(x, x')_{j}$ is sentence pair grouped by 2 user input queries and $l_{j}\in\{0, 1\}$ is similarity class label, see Equ. \ref{eq:similarity label}, 
\begin{align}
l_{j}=\begin{cases} \label{eq:similarity label}
0 & y_{x}\neq y_{x'}\\
1 & y_{x}=y_{x'}
\end{cases}
\end{align}
directly from original data set $\mathcal{D}$ without any sampling strategy, then $\mathcal{D'}$ will have $M=N^2$ data points of sentence pairs, which might be too time-consuming for the training process of stage 2 since we often have hundreds of class labels under task-specific chatbot scenario. Although we may only have a few data points for a single class label, $N$ is still quite a big number.

Therefore, we adopt the idea of adversarial negative sampling strategy raised by \cite{bamler2020extreme}. The key idea of this sampling strategy is to train with negative samples that are hard to be distinguished from positive samples. Now suppose we have a user input query $x_{i}$ and top-K most related class labels set $\mathcal{C'}=\{c'_{1}, \dots, c'_{K}\}$ found by auxiliary model or tool in stage 1, where $K$ is a hyperparameter set by us that controls candidate class number, we apply the adversarial negative sampling strategy into our sentence pair sampling process by the following 2 steps:

\textbf{1. positive sentence pair sampling}: During the training process, we have to first make sure that the ground truth class label $y_{i}$ for $x_{i}$ is within the candidate class label set $\mathcal{C'}$. If $y_{i}\notin \mathcal{C'}$, we will manually add $y_{i}$ into $\mathcal{C'}$ by replacing $c'_{K}$ with $y_{i}$. Afterwards, we will randomly sample $P$ sentences $\{x'_{0},x'_{1}, \dots, x'_{P}\}$ with the same class label as $y_{i}$ from data set $\mathcal{D}$ to form the set of sentence pairs with positive label $\mathcal{P'}_{i}=\{[(x_{i}, x'_{0}), 1], [(x_{i}, x'_{1}), 1], \dots, [(x_{i}, x'_{P}), 1]\}$, where $P$ is also a hyperparameter set by us that controls the number of sentences we should sample from each class.

\textbf{2. negative sentence pair sampling}: As for negative sentence pairs, we will also randomly sample $P$ sentences for each class in the negative candidate class label set $\mathcal{C'}\backslash \{y_{i}\}$. The class labels in $\mathcal{C'}\backslash \{y_{i}\}$ are the set of most confusing classes comparing to the ground truth ${y_{i}}$, so we assume that the sentence pairs with negative label grouped by user input query $x_{i}$ and sentences with class labels in $\mathcal{C'}\backslash \{y_{i}\}$ are also strong adversarial negative samples for sentence-pair similarity model that can help enhance the training speed and performance. According to the same method as positive sentence pair sampling, we can obtain the set of sentence pairs with negative label as $\mathcal{N'}_{i}=\{[(x_{i}, x'_{0}), 0], [(x_{i}, x'_{1}), 0], ..., [(x_{i}, x'_{P\cdot (K-1)}), 0]\}$

@@

\subsubsection{stage 1: classification model for providing top-K classes}
As for stage 1, our SFC system can adapt to any auxiliary model or even searching tool based on Term Frequency-Inverse
Document Frequency (TF-IDF) retrieval. In this paper, however, we choose to fine-tune RoBERTa on our task-specific data set $\mathcal{D}$ by adding a softmax classifier on top of it, the same as the structure shown in Fig. \ref{Fig.1.sub.1}, for better overall performance. According to our experimental result, RoBERTa can guarantee that approximately 95\% ground truth label be included in top-5 candidate classes with highest probability after being fine-tuned on task-specific data set with experimental result measured by F-score at around 0.75. So we use RoBERTa as encoder of classification model in our setting at stage 1 to ensure low limitation for the performance of stage 2.

@@

Before we start to fine-tune our sentence-pair model on task-specific dataset, we first fine-tune RoBERTa on Quora dataset\cite{iyer2017first}, which contains 404,290 potential duplicate question pairs, for transfer learning. Then we continue to do multi-task training on sentence-pair dataset $\mathcal{D'}$ sampled from the top-K candidate classes provided by stage 1.

@@

As for top-K classification task, we share the same sentence-pair similarity model structure with task 1. The only difference is that we add a task-specific average pooling layer, shown in Fig. \ref{fig:ver.1.3} to implement the classification task based on sentence-pair model.

We already know that the size of ${\Phi}^S$ in task 1 is $M\times 2=(N\cdot K\cdot P)\times 2$, where $N$ is the total number of data points in original training data $\mathcal{D}$, $K$ is a hyperparameter that controls the number of candidate classes provided by stage 1, and $P$ is also a hyperparameter that controls the number of sentences we should randomly sample from each candidate class. Now, for the average pooling layer, we first reshape ${\Phi}^S$ into the size of $N\times K\times P\times 2$, and then split it into ${\Phi}^{K,pos}$ and ${\Phi}^{K,neg}$, and both of them will have the size of $N\times K\times P$. In this way, ${\Phi}^{K,pos}$ can represent the level of similarity for each sentence pair, and in the mean time, ${\Phi}^{K,neg}$ can represent the level of dissimilarity for each sentence pair. Then, we can do average pooling for each candidate class among the top-K classes as shown in Equ. \ref{eq:average pooling:1} and Equ. \ref{eq:average pooling:2}.

@@

Therefore, to further improve the overall performance of SFC, we proposed a hierarchical joint model structure, shown in Fig. \ref{fig:framework}. In hierarchical SFC, classification model is being placed in lower layer level to dynamically provide top-K candidate classes for the sentence-pair similarity model placed in higher layer level through a top-K pooling layer. In this way, classification model and similarity model can be merged into one single joint-model for multi-task training, thus avoiding the limitation brought by 2 separate stage structure.

@@

\section{Experiments}
In this section, we conduct extensive experiments over 5 task-specific dataset to evaluate the performance of our proposed SFC system and the effectiveness of the methodology of multi-task learning and hierachical joint-model structure under low-data settings.

\subsection{Datasets}
The dataset used in our experiments aims to help us study low-resource text classification problem under task-specific conversational chatbot scenario. Therefore, we keep all the dataset under low-resource chatbot corpus setting, which is a large number of class labels (50-250) with a few examples(5-20) in short text for each class, by sampling smaller subsets from the full scale when the raw datasets are oversized, see Table \ref{tbe:dataset statistic}.

@@

$\bullet$ \textbf{full-scale real-world task-specific faq chatbot dataset}

@@

$\bullet$ \textbf{subset of customer review dataset for extreme classification}

@@

$\bullet$ \textbf{subset of recently-released intent detection datasets}.

@@

@@

ALBERT-xxlarge & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{$0.1422$} & \multirow{2}*{$0.2057$} & \multirow{2}*{$0.5723$} & \multirow{2}*{$0.5912$} & \multirow{2}*{$0.6601$} & \multirow{2}*{$0.6893$}\tabularnewline

@@

RoBERTa-large & \multirow{2}*{$0.8266$} & \multirow{2}*{$0.8861$} & \multirow{2}*{$0.9023$} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{$0.5425$} & \multirow{2}*{$0.6164$} & \multirow{2}*{$0.6503$} & \multirow{2}*{$0.6729$} & \multirow{2}*{$0.7418$} & \multirow{2}*{$0.6362$}\tabularnewline

@@

2-stage SFC & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-} & \multirow{2}*{-}\tabularnewline

@@

\textbf{Group 1: single-task classification model.} This group of baseline models are based on the network structure for classification model shown in \ref{Fig.1.sub.1}. As for the pretrained model, we choose RoBERTa, BERT and ALBERT, all of which are among the most powerful pretrained models that achieve outstanding results in many NLP tasks, in our experimental setting. Since we apply RoBERTa in our SFC system, we naturally adopt single-task classification model based on RoBERTa-large and RoBERTa-base as our baseline models. Besides, we also adopt the baseline pre-trained model BERT-large\cite{devlin2018bert} and recently released ALBERT-xxlarge\cite{lan2019albert} for comparison. Since we use different few-shot settings from \cite{casanueva2020efficient} for those intent detection datasets, we also use BERT-large tuned on our few-shot settings as an essential reference for experimental results.

@@

\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
\multicolumn{1}{|c|}{\multirow{2}*{Dataset}}&\multicolumn{1}{c|}{\multirow{2}*{Model}}& $K=3$ & $K=5$ & $K=10$ & $K=15$ & $K=20$\tabularnewline
\multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & $P=20$ & $P=10$ & $P=5$ & $P=4$ & $P=3$\tabularnewline
\hline
\multicolumn{1}{|c|}{\multirow{2}*{ITG}}& 2-stage SFC & 0.05 & $0.8124$ & 0.05 & 0.05 & 0.05\tabularnewline
\cline{2-7}
\multicolumn{1}{|c|}{} & Hierarchical SFC & 0.05 & $0.8114$ & 0.05 & 0.05 & 0.05\tabularnewline
\hline
\multicolumn{1}{|c|}{\multirow{2}*{Amazon-670k}}& 2-stage SFC & 0.05 & $0.7364$ & 0.05 & 0.05 & 0.05\tabularnewline
\cline{2-7}
\multicolumn{1}{|c|}{} & Hierarchical SFC & 0.05 & $0.7445$ & 0.05 & 0.05 & 0.05\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\caption{We show the performance of SFC from different settings of hyperparameters for $K$, the candidate class number, and $P$, the sample number for each class.}

@@

First, we turn to compare among the four SFC variants. We conduct ablation experiments to study the benefits brought by multi-task training by comparing multi-task 2-stage SFC with single-task 2-stage SFC. We can see from Table \ref{tbe:table2} that 2-stage SFC with only task 1 (sentence-pair similarity task) degrades by 0.65 percent on average and 2-stage SFC with only task 2 (top-K classification task) degrades by 0.3 percent on average. The degradation indicates that training with these 2 tasks together in SFC can effectively help the sentence-pair model extract more knowledge from the limited amount of data. Besides, task 2 degrades 0.35 percent less than task 1. This observation suggests that task 2 may play a relatively more important role in multi-task learning, which leads us to put larger weight $\alpha_{K}$ to the loss $\mathcal{L}^{K}$ of task 2 when calculating the total loss $\mathcal{L}$. Also, we can see that the hierachical SFC steadily outperforms 2-stage SFC on 5 diverse datasets by 2 percents on average. This observation supports the idea that the hierachical joint model structure can alleviate the limitation brought by the lack of interaction between similarity model and classification model in 2-stage SFC. Therefore, for the following analysis, we will focus on the comparison between hierachical SFC and the other baseline models.

Next, we analyze the overall improvement trend of SFC under various few-shot settings. Our main focus is to study chatbot building in common situation during real-world application where only a few sample sentences are available for each class. The experimental results for the 3 diverse intent detection datasets under few-shot settings indicate that our proposed SFC can achieve +2.47 percents average improvement over one of the most powerful baseline models, which is RoBERTa-large classification model, on all these few-shot settings. Moreover, the improvement in F-score becomes more and more prominent with the decrease of data sample number per class. Our hierarchical SFC can achieve +4.97 percents average improvements over RoBERTa-large classification model in the most extreme few-shot setting with only 5 training samples per class.

Then, we begin to compare the general performance of SFC with classification model and similarity model baselines. As for the classification model, hierarchical SFC outperforms BERT-large by +9.81 percents, ALBERT-xxlarge by 4 percents, RoBERTa-large by +2.47 percents, RoBERTa-base by +4.52 percents on average F-score for 5 datasets. In this way, we can see that RoBERTa is the best pre-trained model for the low-resource text classification task we are studying. As for the similarity model, we only choose the strongest pretrained model RoBERTa-large as our baseline, and hierarchical SFC achieves the improvement of 3 percents on average. Moreover, hierarchical SFC's result on 2 datasets without few-shot settings, which are ITG and Amazon-670k, are also 1.24 percents and 2.89 percents higher than Roberta-large classification model. These improvements illustrate that putting classification model and similarity model at different network layers to cooperate with each other is a successful design of hierarchical network structure for joint model, which enables these two models assist each other for better overall results.

Finally, we analyze the effect brought by different settings of hyperparameters, see Table \ref{tbe:table3}. We conduct experiments on 5 different settings of hyperparameters, which are the number of candidate class number $K$ and the number of sentence pairs $P$ for each class, on ITG and Amazon-670K datasets to find the most suitable setting of these 2 hyperparameters. We control the total number of sentence pairs sampled for each user query into about the same level of amount(50-60). Through analyzing the experimenal results shown in Table \ref{tbe:table3}, we find that the best setting of hyperparameters might be slightly different for datasets of different data size. The class label number of ITG and Amazon-670k are similar in our experimental setting, which is around 250. However the total amount of data points of ITG is approximately 2 times larger than Amazon-670k. Therefore, the setting of $K=5$ and $P=10$ is good enough for ITG since most of the classes contains over 10 data points. However, for Amazon-670k, the amount of data sample for each class is scarce, thus, we may need more candidate classes to sample enough effective sentence pairs for sentence-pair model layer of SFC. In this way, the setting of $K=10$ and $P=5$ becomes a good choice.
